---
title: "Using ggiraph for interactive plot in report"
author: "Federico Tiberti"
date: '2022-04-23'
lastMod: '2022-04-23'
slug: ggiraph-interactive-plot
categories: R
tags:
- R Markdown
- plot
- regression
---

For those looking for an alternative to plotly to create vectorized/interactive graphics keeping ggplot syntax: [ggiraph](https://davidgohel.github.io/ggiraph/#:~:text=ggiraph%20is%20a%20tool%20that,when%20used%20in%20shiny%20applications.) seems to be a great option. Brief example below. I downloaded [some data on wine consumption per capita](https://ourworldindata.org/grapher/wine-consumption-per-person) and joined it to data on GDP per capita, using the great [WDI](https://cran.r-project.org/web/packages/WDI/WDI.pdf) package to interact with the World Bank API (inspired by [a tweet by Andres Lopez](https://twitter.com/anlopez1962/status/1517211524615618560?s=20&t=h4b1KpWNbwA3bTkg8q1Jxw)). I use the awesome [ggflags package](https://github.com/jimjam-slam/ggflags) to plot each country with its flag instead of a point.

```{r, echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

lapply(c("tidyverse", "WDI", "countrycode", "ggiraph", "ggflags"), require, character.only=T,verbose)

gdppc <- WDI::WDI(indicator = "NY.GDP.PCAP.PP.KD", start = 2019, end = 2019) %>% 
  setNames(c("code", "country", "gdppc", "year")) %>% 
  select(-year) %>%
  mutate(code = tolower(code))

vino <- read_csv("wine-consumption-per-person.csv") %>% 
  setNames(c("country", "code", "year", "wine")) %>% 
  filter(year==2019) %>% 
  mutate(code = tolower(countrycode(tolower(code), origin="iso3c", destination = "iso2c"))) %>% 
  left_join(gdppc, by="code")

vino %>% 
  mutate(log_gdp = log(gdppc), 
         log_wine = log(wine)) %>% 
  ggplot(aes(x=log_gdp, y=wine)) +
 theme_classic() +
  geom_flag(aes(country=code), size=3) +
  labs(x="Log GDP per capita, PPP, 2019", y="Wine consumption per capita, 2019") +
  theme(axis.text = element_text(face="bold", size=8), 
        axis.title = element_text(face= "bold", size=12), 
        plot.caption = element_text(size=6),
        plot.title = element_text(size=18, face="bold"))

```


ggiraph plots build on ggplot syntax. The geometry geom_point_interactive is an extended version of geom_point with additional interactive aesthetics.


```{r}
giraffe_points <- vino %>% 
  mutate(country.x = str_replace_all(country.x, "'", ""),
    log_gdp = log(gdppc), 
         log_wine = log(wine)) %>% 
  ggplot(aes(x=log_gdp, y=wine)) +
 theme_classic() +
  geom_point_interactive(aes(tooltip = country.x, data_id = country.x)) +
  labs(x="Log GDP per capita, PPP, 2019", y="Wine consumption per capita, 2019") +
  theme(axis.text = element_text(face="bold", size=8), 
        axis.title = element_text(face= "bold", size=12), 
        plot.caption = element_text(size=6),
        plot.title = element_text(size=18, face="bold"))

girafe(ggobj = giraffe_points)
  
```

Using flags *and* points:

```{r}
giraffe_points <- vino %>% 
  mutate(country.x = str_replace_all(country.x, "'", ""),
    log_gdp = log(gdppc), 
         log_wine = log(wine)) %>% 
  ggplot(aes(x=log_gdp, y=wine)) +
 theme_classic() +
  geom_flag(aes(country=code), size=3) +
  geom_point_interactive(aes(tooltip = country.x, data_id = country.x), alpha=0) +
  labs(x="Log GDP per capita, PPP, 2019", y="Wine consumption per capita, 2019") +
  theme(axis.text = element_text(face="bold", size=8), 
        axis.title = element_text(face= "bold", size=12), 
        plot.caption = element_text(size=6),
        plot.title = element_text(size=18, face="bold"))

girafe(ggobj = giraffe_points)
  
```






